---
title: "Biochemical Plots - Linear Model & Scatterplots"
author: "Juliana Cornett"
date: "1/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Script to plot all biochemical parameters measured (Lactate, SOD, and HIF) in same plot with shared x-axis (DO)


#install packages

```{r}
#install.packages("tidyr")
#install.packages("ggplot2")
#install.packages("tidyverse")
#install.packages("ggpubr")
#install.packages("patchwork")
```

#load libraries

```{r}
library(tidyr)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(patchwork)
```

#read in csv file containing all measurements for each species

```{r}
soleAll <- read.csv("Englishsole_AllVars.csv")

dabAll <- read.csv("Speckledsanddab_AllVars.csv")
```

#select only ID, Species, DO, Lactate, SOD, and HIF columns for each

```{r}
soleBiochem <- select(soleAll, c("ID", "Species", "DO", "Lactate", "SOD", "HIF"))

dabBiochem <- select(dabAll, c("ID", "Species", "DO", "Lactate", "SOD", "HIF"))
```

#SCATTERPLOTS

#ENGLISH SOLE

#linear models

#hif

```{r}
sole_hif <- lm(soleBiochem$HIF ~ soleBiochem$DO)
print(summary(sole_hif))
```

#lactate

```{r}
sole_lac <- lm(soleBiochem$Lactate ~ soleBiochem$DO)
print(summary(sole_lac))
```

#sod

```{r}
sole_sod <- lm(soleBiochem$SOD ~ soleBiochem$DO)
print(summary(sole_sod))
```

#scatterplots combined with patchwork (Figure 12 in thesis)

```{r}
sole_hif_scatter_p1 <- ggplot(soleBiochem, aes(x = DO, y = HIF)) +
  geom_point(fill = "lightskyblue1", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "*`,`~")), size = 6, label.x = 6.1) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("HIF (pg/mL)")

sole_lac_scatter_p2 <- ggplot(soleBiochem, aes(x = DO, y = Lactate)) +
  geom_point(fill = "lightskyblue1", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "*`,`~")), size = 6, label.x = 6.1) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("Lactate (uM)")
  
sole_sod_scatter_p3 <- ggplot(soleBiochem, aes(x = DO, y = SOD)) +
  geom_point(fill = "lightskyblue1", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "*`,`~")), size = 6, label.x = 6.1) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("SOD (U/mL)")

sole_hif_scatter_p1 / sole_lac_scatter_p2 / sole_sod_scatter_p3
```

#SPECKLED SANDDABS

#linear models

#hif

```{r}
dab_hif <- lm(dabBiochem$HIF ~ dabBiochem$DO)
print(summary(dab_hif))
```

#lactate

```{r}
dab_lac <- lm(dabBiochem$Lactate ~ dabBiochem$DO)
print(summary(dab_lac))
```

#sod

```{r}
dab_sod <- lm(dabBiochem$SOD ~ dabBiochem$DO)
print(summary(dab_sod))
```

#scatterplots combined with patchwork (Figure 13 in thesis)

```{r}
dab_hif_scatter_p1 <- ggplot(dabBiochem, aes(x = DO, y = HIF)) +
  geom_point(fill = "lightgoldenrod2", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "*`,`~")), size = 6, label.x = 6.1) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("HIF (pg/mL)")

dab_lac_scatter_p2 <- ggplot(dabBiochem, aes(x = DO, y = Lactate)) +
  geom_point(fill = "lightgoldenrod2", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "*`,`~")), size = 6, label.x = 6.1) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("Lactate (uM)")
  
dab_sod_scatter_p3 <- ggplot(dabBiochem, aes(x = DO, y = SOD)) +
  geom_point(fill = "lightgoldenrod2", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "*`,`~")), size = 6, label.x = 6.1) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("SOD (U/mL)")

dab_hif_scatter_p1 / dab_lac_scatter_p2 / dab_sod_scatter_p3
```
