---
title: "Biochemical Plots - ANOVA, Tukey HSD, & Boxplots"
author: "Juliana Cornett"
date: "3/27/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#install packages

```{r}
#install.packages("tidyr")
#install.packages("ggplot2")
#install.packages("tidyverse")
#install.packages("ggpubr")
#install.packages("patchwork")
#install.packages("multcompView")
```

#load libraries

```{r}
library(tidyr)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(patchwork)
library(multcompView)
```

#DATA PREPARATION

#read in csv file containing all measurements for each species

```{r}
soleAll <- read.csv("Englishsole_AllVars.csv")

dabAll <- read.csv("Speckledsanddab_AllVars.csv")
```

#HIF BOXPLOTS

#ENGLISH SOLE 

#select only DO & HIF

```{r}
soleHIF <- select(soleAll, c("DO", "HIF"))
```

#make DO a factor

```{r}
soleHIF$DO <- as.factor(soleHIF$DO)
```

#remove NAs

```{r}
soleHIF <- na.omit(soleHIF)
```

#ANOVA

```{r}
sole_hif_anova <- aov(HIF ~ DO, data = soleHIF)
summary(sole_hif_anova)
```

#boxplot

```{r}
sole_hif_box <- ggplot(soleHIF, aes(DO, HIF)) + 
  geom_boxplot(fill = "lightskyblue1") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(x=5.5, y=125, label="ANOVA, p = 0.256", size=5) +
  ylab("HIF (pg/mL)") + 
  xlab("DO (mg/L)")
```

#SPECKLED SANDDABS

#select only DO & HIF

```{r}
dabHIF <- select(dabAll, c("DO", "HIF"))
```

#make DO a factor

```{r}
dabHIF$DO <- as.factor(dabHIF$DO)
```

#remove NAs

```{r}
dabHIF <- na.omit(dabHIF)
```

#ANOVA

```{r}
dab_hif_anova <- aov(HIF ~ DO, data = dabHIF)
summary(dab_hif_anova)
```

#boxplot

```{r}
dab_hif_box <- ggplot(dabHIF, aes(DO, HIF)) + 
  geom_boxplot(fill = "lightgoldenrod2") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(x=5.5, y=97, label="ANOVA, p = 0.174", size=5) +
  ylab("HIF (pg/mL)") + 
  xlab("DO (mg/L)")
```

#LACTATE BOXPLOTS

#ENGLISH SOLE 

#select only DO & Lactate

```{r}
soleLac <- select(soleAll, c("DO", "Lactate"))
```

#make DO a factor

```{r}
soleLac$DO <- as.factor(soleLac$DO)
```

#remove NAs

```{r}
soleLac <- na.omit(soleLac)
```

#ANOVA

```{r}
sole_lac_anova <- aov(Lactate ~ DO, data = soleLac)
summary(sole_lac_anova)
```

#Tukey HSD post-hoc test

```{r}
sole_lac_tukey <- TukeyHSD(sole_lac_anova)
print(sole_lac_tukey)
```

#compact letter display of Tukey HSD results

```{r}
sole_lac_cld <- multcompLetters4(sole_lac_anova, sole_lac_tukey)
print(sole_lac_cld)
```

#make a dataframe

```{r}
# table with factors and 3rd quantile
sole_lac_table <- group_by(soleLac, DO) %>%
  summarise(mean=mean(Lactate), quant = quantile(Lactate, probs = 0.75)) %>%
  arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
sole_lac_cld <- as.data.frame.list(sole_lac_cld$DO)
sole_lac_table$sole_lac_cld <- sole_lac_cld$Letters

print(sole_lac_table)
```

#boxplot

```{r}
sole_lac_box <- ggplot(soleLac, aes(DO, Lactate)) + 
  geom_boxplot(fill = "lightskyblue1") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(data = sole_lac_table, aes(x = DO, y = quant, label = sole_lac_cld, vjust=-1, hjust =-1)) +
  geom_text(x=5.5, y=2500, label="ANOVA, p < 0.001", size=5) +
  ylab("Lactate (uM)") + 
  xlab("DO (mg/L)")
```

#SPECKLED SANDDABS

#select only DO & Lactate

```{r}
dabLac <- select(dabAll, c("DO", "Lactate"))
```

#make DO a factor

```{r}
dabLac$DO <- as.factor(dabLac$DO)
```

#remove NAs

```{r}
dabLac <- na.omit(dabLac)
```

#ANOVA

```{r}
dab_lac_anova <- aov(Lactate ~ DO, data = dabLac)
summary(dab_lac_anova)
```

#Tukey HSD post-hoc test

```{r}
dab_lac_tukey <- TukeyHSD(dab_lac_anova)
print(dab_lac_tukey)
```

#compact letter display of Tukey HSD results

```{r}
dab_lac_cld <- multcompLetters4(dab_lac_anova, dab_lac_tukey)
print(dab_lac_cld)
```

#make a dataframe

```{r}
# table with factors and 3rd quantile
dab_lac_table <- group_by(dabLac, DO) %>%
  summarise(mean=mean(Lactate), quant = quantile(Lactate, probs = 0.75)) %>%
  arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
dab_lac_cld <- as.data.frame.list(dab_lac_cld$DO)
dab_lac_table$dab_lac_cld <- dab_lac_cld$Letters

print(dab_lac_table)
```

#boxplot

```{r}
dab_lac_box <- ggplot(dabLac, aes(DO, Lactate)) + 
  geom_boxplot(fill = "lightgoldenrod2") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(data = dab_lac_table, aes(x = DO, y = quant, label = dab_lac_cld, vjust=-1, hjust =-1)) +
  geom_text(x=5.5, y=5250, label="ANOVA, p = 0.0123", size=5) +
  ylab("Lactate (uM)") + 
  xlab("DO (mg/L)")
```

#SOD BOXPLOTS

#ENGLISH SOLE 

#select only DO & SOD

```{r}
soleSOD <- select(soleAll, c("DO", "SOD"))
```

#make DO a factor

```{r}
soleSOD$DO <- as.factor(soleSOD$DO)
```

#remove NAs

```{r}
soleSOD <- na.omit(soleSOD)
```

#ANOVA

```{r}
sole_sod_anova <- aov(SOD ~ DO, data = soleSOD)
summary(sole_sod_anova)
```

#Tukey HSD post-hoc test

```{r}
sole_sod_tukey <- TukeyHSD(sole_sod_anova)
print(sole_sod_tukey)
```

#compact letter display of Tukey HSD results

```{r}
sole_sod_cld <- multcompLetters4(sole_sod_anova, sole_sod_tukey)
print(sole_sod_cld)
```

#make a dataframe

```{r}
# table with factors and 3rd quantile
sole_sod_table <- group_by(soleSOD, DO) %>%
  summarise(mean=mean(SOD), quant = quantile(SOD, probs = 0.75)) %>%
  arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
sole_sod_cld <- as.data.frame.list(sole_sod_cld$DO)
sole_sod_table$sole_sod_cld <- sole_sod_cld$Letters

print(sole_sod_table)
```

#boxplot

```{r}
sole_sod_box <- ggplot(soleSOD, aes(DO, SOD)) + 
  geom_boxplot(fill = "lightskyblue1") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(data = sole_sod_table, aes(x = DO, y = quant, label = sole_sod_cld, vjust=-1, hjust =-1)) +
  geom_text(x=5.5, y=2.75, label="ANOVA, p = 0.0434", size=5) +
  ylab("SOD (U/mL)") + 
  xlab("DO (mg/L)")
```

#SPECKLED SANDDABS

#select only DO & SOD

```{r}
dabSOD <- select(dabAll, c("DO", "SOD"))
```

#make DO a factor

```{r}
dabSOD$DO <- as.factor(dabSOD$DO)
```

#remove NAs

```{r}
dabSOD <- na.omit(dabSOD)
```

#ANOVA

```{r}
dab_sod_anova <- aov(SOD ~ DO, data = dabSOD)
summary(dab_sod_anova)
```

#boxplot

```{r}
dab_sod_box <- ggplot(dabSOD, aes(DO, SOD)) + 
  geom_boxplot(fill = "lightgoldenrod2") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(x=5.5, y=9.1, label="ANOVA, p = 0.277", size=5) +
  ylab("SOD (U/mL)") + 
  xlab("DO (mg/L)")
```

#JOIN PLOTS with PATCHWORK

#ENGLISH SOLE (Figure 10 in thesis)

```{r}
sole_hif_box / sole_lac_box / sole_sod_box
```

#SPECKLED SANDDABS (Figure 11 in thesis)

```{r}
dab_hif_box / dab_lac_box / dab_sod_box
```

#ANOVAs for Species Comparison Table p-values in Thesis (Table 4)

#dataframe with both species

```{r}
BothBiochem <- rbind(soleAll, dabAll)
```

#anovas of each metric as a function of DO and Species + interaction

```{r}
aov_lac <- aov(Lactate ~ as.factor(DO) * as.factor(Species), data = BothBiochem)
summary(aov_lac)
```

```{r}
aov_hif <- aov(HIF ~ as.factor(DO) * as.factor(Species), data = BothBiochem)
summary(aov_hif)
```

```{r}
aov_sod <- aov(SOD ~ as.factor(DO) * as.factor(Species), data = BothBiochem)
summary(aov_sod)
```
