---
title: "Compensatory Mechanism Plots - ANOVA, Tukey HSD, & Boxplots"
author: "Juliana Cornett"
date: "3/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#install packages

```{r}
#install.packages("tidyr")
#install.packages("ggplot2")
#install.packages("tidyverse")
#install.packages("ggpubr")
#install.packages("patchwork")
#install.packages("multcompView")
```

#load libraries

```{r}
library(tidyr)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(patchwork)
library(multcompView)
```

#DATA PREPARATION

#read in csv file containing all measurements for each species

```{r}
soleAll <- read.csv("Englishsole_AllVars.csv")

dabAll <- read.csv("Speckledsanddab_AllVars.csv")
```

#select only ID, Species, DO, Lactate, SOD, and HIF columns for each

```{r}
soleComp <- select(soleAll, c("ID", "Species", "DO", "Hct", "AmbVent", "HypVent"))

dabComp <- select(dabAll, c("ID", "Species", "DO", "Hct", "AmbVent", "HypVent"))
```

#create ventilation rate change column (hypVent - AmbVent)

```{r}
soleComp$VentChange <- (soleComp$HypVent - soleComp$AmbVent)

dabComp$VentChange <- (dabComp$HypVent - dabComp$AmbVent)
```

#VENTILATION BOXPLOTS

#ENGLISH SOLE

#select only DO & VentChange

```{r}
soleVent <- select(soleComp, c("DO", "VentChange"))
```

#make DO a factor

```{r}
soleVent$DO <- as.factor(soleVent$DO)
```

#remove NAs

```{r}
soleVent <- na.omit(soleVent)
```

#ANOVA

```{r}
sole_vent_anova <- aov(VentChange ~ DO, data = soleVent)
summary(sole_vent_anova)
```

#Tukey HSD post-hoc test

```{r}
sole_vent_tukey <- TukeyHSD(sole_vent_anova)
print(sole_vent_tukey)
```

#compact letter display of Tukey HSD results

```{r}
sole_vent_cld <- multcompLetters4(sole_vent_anova, sole_vent_tukey)
print(sole_vent_cld)
```

#make a dataframe

```{r}
# table with factors and 3rd quantile
sole_vent_table <- group_by(soleVent, DO) %>%
  summarise(mean=mean(VentChange), quant = quantile(VentChange, probs = 0.75)) %>%
  arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
sole_vent_cld <- as.data.frame.list(sole_vent_cld$DO)
sole_vent_table$sole_vent_cld <- sole_vent_cld$Letters

print(sole_vent_table)
```

#boxplot

```{r}
sole_vent_box <- ggplot(soleVent, aes(DO, VentChange)) + 
  geom_boxplot(fill = "lightskyblue1") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(data = sole_vent_table, aes(x = DO, y = quant, label = sole_vent_cld, vjust=-1, hjust =-1)) +
  geom_text(x=1.5, y=0, label="ANOVA, p = 0.00507", size=5) +
  ylab("Change in Ventilation Rate (breaths/min)") + 
  xlab("DO (mg/L)")
```

#SPECKLED SANDDABS

#select only DO & VentChange

```{r}
dabVent <- select(dabComp, c("DO", "VentChange"))
```

#make DO a factor

```{r}
dabVent$DO <- as.factor(dabVent$DO)
```

#remove NAs

```{r}
dabVent <- na.omit(dabVent)
```

#ANOVA

```{r}
dab_vent_anova <- aov(VentChange ~ DO, data = dabVent)
summary(dab_vent_anova)
```

#Tukey HSD post-hoc test

```{r}
dab_vent_tukey <- TukeyHSD(dab_vent_anova)
print(dab_vent_tukey)
```

#compact letter display of Tukey HSD results

```{r}
dab_vent_cld <- multcompLetters4(dab_vent_anova, dab_vent_tukey)
print(dab_vent_cld)
```

#make a dataframe

```{r}
# table with factors and 3rd quantile
dab_vent_table <- group_by(dabVent, DO) %>%
  summarise(mean=mean(VentChange), quant = quantile(VentChange, probs = 0.75)) %>%
  arrange(desc(mean))

# extracting the compact letter display and adding to the table
dab_vent_cld <- as.data.frame.list(dab_vent_cld$DO)
dab_vent_table$dab_vent_cld <- dab_vent_cld$Letters

print(dab_vent_table)
```

#boxplot

```{r}
dab_vent_box <- ggplot(dabVent, aes(DO, VentChange)) + 
  geom_boxplot(fill = "lightgoldenrod2") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(data = dab_vent_table, aes(x = DO, y = quant, label = dab_vent_cld, vjust=-1, hjust =-1)) +
  geom_text(x=1.5, y=0, label="ANOVA, p < 0.001", size=5) +
  ylab("Change in Ventilation Rate (breaths/min)") + 
  xlab("DO (mg/L)")
```

#HEMATOCRIT BOXPLOTS

#ENGLISH SOLE

#select only DO & Hct

```{r}
soleHct <- select(soleComp, c("DO", "Hct"))
```

#make DO a factor

```{r}
soleHct$DO <- as.factor(soleHct$DO)
```

#remove NAs

```{r}
soleHct <- na.omit(soleHct)
```

#ANOVA

```{r}
sole_hct_anova <- aov(Hct ~ DO, data = soleHct)
summary(sole_hct_anova)
```

#Tukey HSD post-hoc test

```{r}
sole_hct_tukey <- TukeyHSD(sole_hct_anova)
print(sole_hct_tukey)
```

#compact letter display of Tukey HSD results

```{r}
sole_hct_cld <- multcompLetters4(sole_hct_anova, sole_hct_tukey)
print(sole_hct_cld)
```

#make a dataframe

```{r}
# table with factors and 3rd quantile
sole_hct_table <- group_by(soleHct, DO) %>%
  summarise(mean=mean(Hct), quant = quantile(Hct, probs = 0.75)) %>%
  arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
sole_hct_cld <- as.data.frame.list(sole_hct_cld$DO)
sole_hct_table$sole_hct_cld <- sole_hct_cld$Letters

print(sole_hct_table)
```

#boxplot

```{r}
sole_hct_box <- ggplot(soleHct, aes(DO, Hct)) + 
  geom_boxplot(fill = "lightskyblue1") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(data = sole_hct_table, aes(x = DO, y = quant, label = sole_hct_cld, vjust=-1, hjust =-1)) +
  geom_text(x=1.5, y=20, label="ANOVA, p = 0.0502", size=5) +
  ylab("Hematocrit (%)") + 
  xlab("DO (mg/L)")
```


#SPECKLED SANDDABS

#select only DO & Hct

```{r}
dabHct <- select(dabComp, c("DO", "Hct"))
```

#make DO a factor

```{r}
dabHct$DO <- as.factor(dabHct$DO)
```

#remove NAs

```{r}
dabHct <- na.omit(dabHct)
```

#ANOVA

```{r}
dab_hct_anova <- aov(Hct ~ DO, data = dabHct)
summary(dab_hct_anova)
```

#Tukey HSD post-hoc test

```{r}
dab_hct_tukey <- TukeyHSD(dab_hct_anova)
print(dab_hct_tukey)
```

#compact letter display of Tukey HSD results

```{r}
dab_hct_cld <- multcompLetters4(dab_hct_anova, dab_hct_tukey)
print(dab_hct_cld)
```

#make a dataframe

```{r}
# table with factors and 3rd quantile
dab_hct_table <- group_by(dabHct, DO) %>%
  summarise(mean=mean(Hct), quant = quantile(Hct, probs = 0.75)) %>%
  arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
dab_hct_cld <- as.data.frame.list(dab_hct_cld$DO)
dab_hct_table$dab_hct_cld <- dab_hct_cld$Letters

print(dab_hct_table)
```

#boxplot

```{r}
dab_hct_box <- ggplot(dabHct, aes(DO, Hct)) + 
  geom_boxplot(fill = "lightgoldenrod2") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(data = dab_hct_table, aes(x = DO, y = quant, label = dab_hct_cld, vjust=-1, hjust =-1)) +
  geom_text(x=1.5, y=27, label="ANOVA, p = 0.0231", size=5) +
  ylab("Hematocrit (%)") + 
  xlab("DO (mg/L)")
```

#JOIN PLOTS with PATCHWORK

#ENGLISH SOLE

```{r}
sole_hct_box / sole_vent_box
```

#SPECKLED SANDDABS

```{r}
dab_hct_box / dab_vent_box
```

#SPECIES COMPARISON (used in Table 4 in thesis)

#join dataframes

```{r}
bothComp <- rbind(soleComp, dabComp)
```

#create ventilation rate change column

```{r}
bothComp$ventDiff <- (bothComp$HypVent - bothComp$AmbVent)
```

#two-way ANOVA with interaction for Hct

```{r}
Hct_aov_inter <- aov(Hct ~ as.factor(DO) * as.factor(Species), data = bothComp)
summary(Hct_aov_inter)
```

#two-way anova with interaction for ventilation

```{r}
vent_aov_inter <- aov(ventDiff ~ as.factor(DO) * as.factor(Species), data = bothComp)
summary(vent_aov_inter)
```
