---
title: "MetabolicPlots - ANOVA, Tukey HSD, & Boxplots"
author: "Juliana Cornett"
date: "3/27/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#install required packages (tidyverse, ggpubr, multcompView, and pathchwork)

```{r}
#install.packages("tidyverse")
#install.packages("ggpubr")
#install.packages("patchwork")
#install.packages("multcompView")
```

#load libraries

```{r}
library(tidyverse)
library(ggpubr)
```

#read in CSV files with MMR & SMR data for English sole at each DO level

```{r}
sole_scope <- read.csv("../Data/Scope_Pvetulus.csv")
```

#create new ambient and hypoxic aerobic scope columns by subtracting MMR - SMR

```{r}
sole_scope$AmbScope <- (sole_scope$AmbMMR - sole_scope$AmbSMR)
sole_scope$HypScope <- (sole_scope$HypMMR - sole_scope$HypSMR)
```

#create aerobic scope difference column by subtracting hypoxic - ambient scope

```{r}
sole_scope$ScopeChange <- (sole_scope$HypScope - sole_scope$AmbScope)
```

#add MMR change and SMR change columns to scope dataframe

```{r}
sole_scope$MMRChange <- (sole_scope$HypMMR - sole_scope$AmbMMR)
sole_scope$SMRChange <- (sole_scope$HypSMR - sole_scope$AmbSMR)
```

#POST-EXPOSURE BOXPLOTS

#SMR

#select only DO & HypSMR (post-exposure SMR)

```{r}
PostSMR <- select(sole_scope, c("DO_mgL", "HypSMR"))
```

#make DO a factor

```{r}
PostSMR$DO_mgL <- as.factor(PostSMR$DO_mgL)
```

#remove NAs

```{r}
PostSMR <- na.omit(PostSMR)
```

#ANOVA

```{r}
postSMR_anova <- aov(HypSMR ~ DO_mgL, data = PostSMR)
summary(postSMR_anova)
```

#boxplot

```{r}
postSMR_box <- ggplot(PostSMR, aes(DO_mgL, HypSMR)) + 
  geom_boxplot(fill = "lightskyblue1") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(x=1.5, y=600, label="ANOVA, p = 0.893", size=5) +
  ylab("Post-exposure SMR (mgO2/kg/hr)") + 
  xlab("DO (mg/L)")
```

#MMR

#select only DO & HypMMR (post-exposure MMR)

```{r}
PostMMR <- select(sole_scope, c("DO_mgL", "HypMMR"))
```

#make DO a factor

```{r}
PostMMR$DO_mgL <- as.factor(PostMMR$DO_mgL)
```

#remove NAs

```{r}
PostMMR <- na.omit(PostMMR)
```

#ANOVA

```{r}
postMMR_anova <- aov(HypMMR ~ DO_mgL, data = PostMMR)
summary(postMMR_anova)
```

#boxplot

```{r}
postMMR_box <- ggplot(PostMMR, aes(DO_mgL, HypMMR)) + 
  geom_boxplot(fill = "lightskyblue1") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(x=1.5, y=600, label="ANOVA, p = 0.588", size=5) +
  ylab("Post-exposure MMR (mgO2/kg/hr)") + 
  xlab("DO (mg/L)")
```

#AEROBIC SCOPE

#select only DO & HypScope (post-exposure aerobic scope)

```{r}
PostScope <- select(sole_scope, c("DO_mgL", "HypScope"))
```

#make DO a factor

```{r}
PostScope$DO_mgL <- as.factor(PostScope$DO_mgL)
```

#remove NAs

```{r}
PostScope <- na.omit(PostScope)
```

#ANOVA

```{r}
postScope_anova <- aov(HypScope ~ DO_mgL, data = PostScope)
summary(postScope_anova)
```

#boxplot

```{r}
postScope_box <- ggplot(PostScope, aes(DO_mgL, HypScope)) + 
  geom_boxplot(fill = "lightskyblue1") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(x=1.5, y=-50, label="ANOVA, p = 0.0688", size=5) +
  ylab("Post-exposure Aerobic Scope (mgO2/kg/hr)") + 
  xlab("DO (mg/L)")
```

#PRE TO POST-EXPOSURE CHANGE BOXPLOTS

#SMR

#select only DO & SMRChange (pre to post-exposure change in SMR)

```{r}
changeSMR <- select(sole_scope, c("DO_mgL", "SMRChange"))
```

#make DO a factor

```{r}
changeSMR$DO_mgL <- as.factor(changeSMR$DO_mgL)
```

#remove NAs

```{r}
changeSMR <- na.omit(changeSMR)
```

#ANOVA

```{r}
changeSMR_anova <- aov(SMRChange ~ DO_mgL, data = changeSMR)
summary(changeSMR_anova)
```

#Tukey HSD post-hoc test

```{r}
changeSMR_tukey <- TukeyHSD(changeSMR_anova)
print(changeSMR_tukey)
```

#compact letter display of Tukey HSD results

```{r}
changeSMR_cld <- multcompLetters4(changeSMR_anova, changeSMR_tukey)
print(changeSMR_cld)
```

#make a dataframe

```{r}
# table with factors and 3rd quantile
changeSMR_table <- group_by(changeSMR, DO_mgL) %>%
  summarise(mean=mean(SMRChange), quant = quantile(SMRChange, probs = 0.75)) %>%
  arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
changeSMR_cld <- as.data.frame.list(changeSMR_cld$DO_mgL)
changeSMR_table$changeSMR_cld <- changeSMR_cld$Letters

print(changeSMR_table)
```

#boxplot

```{r}
changeSMR_box <- ggplot(changeSMR, aes(DO_mgL, SMRChange)) + 
  geom_boxplot(fill = "lightskyblue1") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(data = changeSMR_table, aes(x = DO_mgL, y = quant, label = changeSMR_cld, vjust=-1, hjust =-1)) +
  geom_text(x=1.5, y=-300, label="ANOVA, p = 0.0366", size=5) +
  ylab("Change in SMR (mgO2/kg/hr)") + 
  xlab("DO (mg/L)")
```

#MMR

#select only DO & MMRChange (pre to post-exposure change in MMR)

```{r}
changeMMR <- select(sole_scope, c("DO_mgL", "MMRChange"))
```

#make DO a factor

```{r}
changeMMR$DO_mgL <- as.factor(changeMMR$DO_mgL)
```

#remove NAs

```{r}
changeMMR <- na.omit(changeMMR)
```

#ANOVA

```{r}
changeMMR_anova <- aov(MMRChange ~ DO_mgL, data = changeMMR)
summary(changeMMR_anova)
```

#Tukey HSD post-hoc test

```{r}
changeMMR_tukey <- TukeyHSD(changeMMR_anova)
print(changeMMR_tukey)
```

#compact letter display of Tukey HSD results

```{r}
changeMMR_cld <- multcompLetters4(changeMMR_anova, changeMMR_tukey)
print(changeMMR_cld)
```

#make a dataframe

```{r}
# table with factors and 3rd quantile
changeMMR_table <- group_by(changeMMR, DO_mgL) %>%
  summarise(mean=mean(MMRChange), quant = quantile(MMRChange, probs = 0.75)) %>%
  arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
changeMMR_cld <- as.data.frame.list(changeMMR_cld$DO_mgL)
changeMMR_table$changeMMR_cld <- changeMMR_cld$Letters

print(changeMMR_table)
```

#boxplot

```{r}
changeMMR_box <- ggplot(changeMMR, aes(DO_mgL, MMRChange)) + 
  geom_boxplot(fill = "lightskyblue1") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(data = changeMMR_table, aes(x = DO_mgL, y = quant, label = changeMMR_cld, vjust=-1, hjust =-1)) +
  geom_text(x=2.25, y=-700, label="ANOVA, p < 0.001", size=5) +
  ylab("Change in MMR (mgO2/kg/hr)") + 
  xlab("DO (mg/L)")
```

#AEROBIC SCOPE

#select only DO & ScopeChange (pre to post-exposure change in Aerobic Scope)

```{r}
changeScope <- select(sole_scope, c("DO_mgL", "ScopeChange"))
```

#make DO a factor

```{r}
changeScope$DO_mgL <- as.factor(changeScope$DO_mgL)
```

#remove NAs

```{r}
changeScope <- na.omit(changeScope)
```

#ANOVA

```{r}
changeScope_anova <- aov(ScopeChange ~ DO_mgL, data = changeScope)
summary(changeScope_anova)
```

#Tukey HSD post-hoc test

```{r}
changeScope_tukey <- TukeyHSD(changeScope_anova)
print(changeScope_tukey)
```

#compact letter display of Tukey HSD results

```{r}
changeScope_cld <- multcompLetters4(changeScope_anova, changeScope_tukey)
print(changeScope_cld)
```

#make a dataframe

```{r}
# table with factors and 3rd quantile
changeScope_table <- group_by(changeScope, DO_mgL) %>%
  summarise(mean=mean(ScopeChange), quant = quantile(ScopeChange, probs = 0.75)) %>%
  arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
changeScope_cld <- as.data.frame.list(changeScope_cld$DO_mgL)
changeScope_table$changeScope_cld <- changeScope_cld$Letters

print(changeScope_table)
```

#boxplot

```{r}
changeScope_box <- ggplot(changeScope, aes(DO_mgL, ScopeChange)) + 
  geom_boxplot(fill = "lightskyblue1") +
  geom_point(shape=1) +
  theme_pubr() +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  geom_text(data = changeScope_table, aes(x = DO_mgL, y = quant, label = changeScope_cld, vjust=-1, hjust =-1)) +
  geom_text(x=2.25, y=-575, label="ANOVA, p = 0.0539", size=5) +
  ylab("Change in Aerobic Scope (mgO2/kg/hr)") + 
  xlab("DO (mg/L)")
```

#BOXPLOTS

```{r} 
jpeg("EnglishSole_MetabolicBoxplots_Patchwork.jpg", width = 14, height = 16, units="in", res=1200)

(postSMR_box | changeSMR_box)/
  (postMMR_box | changeMMR_box)/
  (postScope_box | changeScope_box)

dev.off()
```

