---
title: "Metabolic Plots - Linear Model & Scatterplots"
author: "Juliana Cornett"
date: "1/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#install required packages (tidyverse & ggpubr)

```{r}
#install.packages("tidyverse")
#install.packages("ggpubr")
```

#load libraries

```{r}
library(tidyverse)
library(ggpubr)
```

#DATA PREPARATION

#read in CSV files with MMR & SMR data for English sole at each DO level

```{r}
sole_scope <- read.csv("../Data/Scope_Pvetulus.csv")
```

```{r}
sole_scope
```

#create new ambient and hypoxic aerobic scope columns by subtracting MMR - SMR

```{r}
sole_scope$AmbScope <- (sole_scope$AmbMMR - sole_scope$AmbSMR)
sole_scope$HypScope <- (sole_scope$HypMMR - sole_scope$HypSMR)
```

#create aerobic scope difference column by subtracting hypoxic - ambient scope

```{r}
sole_scope$ScopeChange <- (sole_scope$HypScope - sole_scope$AmbScope)
```

#add MMR change and SMR change columns to scope dataframe

```{r}
sole_scope$MMRChange <- (sole_scope$HypMMR - sole_scope$AmbMMR)
sole_scope$SMRChange <- (sole_scope$HypSMR - sole_scope$AmbSMR)
```

#linear models

#post SMR

```{r}
post_SMR <- lm(sole_scope$HypSMR ~ sole_scope$DO_mgL)
print(summary(post_SMR))
```

#change SMR

```{r}
change_SMR <- lm(sole_scope$SMRChange ~ sole_scope$DO_mgL)
print(summary(change_SMR))
```

#post MMR

```{r}
post_MMR <- lm(sole_scope$HypMMR ~ sole_scope$DO_mgL)
print(summary(post_MMR))
```

#change MMR

```{r}
change_MMR <- lm(sole_scope$MMRChange ~ sole_scope$DO_mgL)
print(summary(change_MMR))
```

#post scope

```{r}
post_scope <- lm(sole_scope$HypScope ~ sole_scope$DO_mgL)
print(summary(post_scope))
```

#change scope

```{r}
change_scope <- lm(sole_scope$ScopeChange ~ sole_scope$DO_mgL)
print(summary(change_scope))
```

#POST-EXPOSURE + CHANGE SCATTERPLOTS with PATCHWORK

```{r}
jpeg("EnglishSole_MetabolicScatterplots_Patchwork.jpg", width = 14, height = 14, units="in", res=1200)

SMR_Post_p1 <- ggplot(sole_scope, aes(x = DO_mgL, y = HypSMR)) +
  geom_point(fill = "lightskyblue1", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "*`,`~")), label.x.npc = "left", label.y.npc = "top", size=6, face="bold") +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=16, face="bold")) + 
  xlab(NULL) +
  ylab("Post-exposure SMR (mgO2/kg/hr)")

SMR_Change_p2 <- ggplot(sole_scope, aes(x = DO_mgL, y = SMRChange)) +
  geom_point(fill = "lightskyblue1", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "*`,`~")), label.x = 5.5, label.y.npc = "bottom", size=6, face="bold") +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=16, face="bold")) +
  geom_hline(yintercept=0, linetype='dotted') + 
  xlab(NULL) +
  ylab("Change in SMR (mgO2/kg/hr)")

MMR_Post_p3 <- ggplot(sole_scope, aes(x = DO_mgL, y = HypMMR)) +
  geom_point(fill = "lightskyblue1", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "*`,`~")), label.x.npc = "left", label.y.npc = "top", size=6, face="bold") +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=16, face="bold")) + 
  xlab(NULL) +
  ylab("Post-exposure MMR (mgO2/kg/hr)")

MMR_Change_p4 <- ggplot(sole_scope, aes(x = DO_mgL, y = MMRChange)) +
  geom_point(fill = "lightskyblue1", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  geom_text(aes(x=6.85, y=-600, label= "R^2 == 0.33~~~p < 0.001"), size=6, parse = TRUE) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=16, face="bold")) +
  geom_hline(yintercept=0, linetype='dotted') + 
  xlab(NULL) +
  ylab("Change in MMR (mgO2/kg/hr)")

Scope_Post_p5 <- ggplot(sole_scope, aes(x = DO_mgL, y = HypScope)) +
  geom_point(fill = "lightskyblue1", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "*`,`~")), label.x.npc = "left", label.y.npc = "top", size=6, face="bold") +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=16, face="bold")) + 
  xlab("DO (mg/L)") +
  ylab("Post-exposure Aerobic Scope (mgO2/kg/hr)")

Scope_Change_p6 <- ggplot(sole_scope, aes(x = DO_mgL, y = ScopeChange)) +
  geom_point(fill = "lightskyblue1", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "*`,`~")), label.x = 5.5, label.y.npc = "bottom", size=6, face="bold") +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=16, face="bold")) +
  geom_hline(yintercept=0, linetype='dotted') + 
  xlab("DO (mg/L)") +
  ylab("Change in Aerobic Scope (mgO2/kg/hr)")

(SMR_Post_p1 | SMR_Change_p2)/
  (MMR_Post_p3 | MMR_Change_p4)/
  (Scope_Post_p5 | Scope_Change_p6)

dev.off()
```
