---
title: "ElkhornSloughDO"
author: "Juliana Cornett"
date: "1/10/2022"
output: html_document
---

#Plot of DO in Elkhorn Slough

#install packages

```{r}
#install.packages("tidyr")
#install.packages("ggplot2")
#install.packages("tidyverse")
#install.packages("ggpubr")
#install.packages("lubridate")
#install.packages("scales")
#install.packages("viridis")
#install.packages("RColorBrewer")
```

#load libraries

```{r}
library(tidyr)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(lubridate)
library(scales)
library(viridis)
library(RColorBrewer)
```

#read in csv files containing DO measurements from 5/1/31-8/31/21 at the 4 water quality sites in Elkhorn Slough (Vierra Mouth, South Marsh, North Marsh, and Azevedo Pond)

```{r}
Vierra <- read.csv("../Data/VierraMouthDO_5.1.21_8.31.21.csv")

South <- read.csv("../Data/SouthMarsh_DO_5.1.21_8.31.21.csv")

North <- read.csv("../Data/NorthMarsh_DO_5.1.21_8.31.21.csv")

Azevedo <- read.csv("../Data/AzevedoPond_DO_5.1.21_8.31.21.csv")
```

#combine into single dataframe

```{r}
AllES_DO <- rbind(Vierra, South, North, Azevedo)
```

#remove rows with NA

```{r}
ES_DO <- na.omit(AllES_DO)
```

#fix timestamp column
#first change slashes in year of TimeStamp column to dashes with gsub

```{r}
ES_DO$TimeStamp <- gsub('/', '-', ES_DO$TimeStamp)
ES_DO
```

#then convert time stamp into POSIXct format, specifying the format that the timestamp is currently in

```{r}
ES_DO$TimeStamp <- as.POSIXct(ES_DO$TimeStamp, format="%m-%d-%Y %H:%M")
ES_DO
```

#plot

```{r}
ggplot(ES_DO, aes(x = TimeStamp, y = DO_mgL, color = Site)) + 
  theme_pubr() + 
  geom_line() +
  ylab("Dissolved Oxygen (mg/L)") + 
  xlab("Date") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme(axis.title=element_text(size=18,face="bold")) +
  theme(axis.text=element_text(size=10)) + 
  theme(legend.text=element_text(size=12,face="bold")) +
  theme(legend.title = element_blank()) 
```

#add month column

```{r}
ES_DO$Month <- as.numeric(format(ES_DO$TimeStamp, "%m"))
ES_DO
```

#subset May data

```{r}
May_DO <- subset(ES_DO, Month==5)
May_DO
```


#set my own color palette

```{r}
palette_DO <- c("#F15946", "#FBB13C", "#3CBBB1", "#004E64")
palette_viridis <- c("#FDE725FF", "#95D840FF", "#238A8DFF", "#482677FF")
palette_v <- c("#482677FF", "#2D7D8EFF", "#55C667FF","#FDE725FF")
```

#plot May only 

```{r}
#jpeg("ElkhornSloughDO.jpg", width = 7, height = 5, units="in", res=1200)
ggplot(May_DO, aes(x = TimeStamp, y = DO_mgL, color = Site)) + 
  scale_color_manual(values = palette_v) +
  theme_pubr() + 
  geom_line() +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ylab("Dissolved Oxygen (mg/L)") + 
  xlab("Date") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme(axis.title=element_text(size=14,face="bold")) +
  theme(axis.text=element_text(size=10)) + 
  theme(legend.text=element_text(size=10,face="bold")) +
  theme(legend.title = element_blank()) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%b %d") + 
  geom_hline(yintercept=8, linetype='dotted')
#dev.off()
```

#add day column to May DO data

```{r}
May_DO$Day <- as.numeric(format(May_DO$TimeStamp, "%d"))
May_DO
```
#add hour column to May DO data

```{r}
May_DO$Hour <- format(May_DO$TimeStamp, format = "%H:%M:%S")
May_DO
```
#subset from 12 pm on May 5th to 12 pm on May 6th

```{r}
May5_6_subset <- subset(May_DO, TimeStamp >= as.POSIXct('0021-05-05 12:00') & TimeStamp <=
as.POSIXct('0021-05-06 12:00'))
May5_7_subset
```


#subset May 6th data only

```{r}
May6_DO <- subset(May_DO, Day==6)
May6_DO
```

#subset May 5th data only

```{r}
May5_DO <- subset(May_DO, Day==5)
May5_DO
```

#plot May 6th data only 

```{r}
jpeg("ElkhornSloughDO_5.6.21.jpg", width = 7, height = 5, units="in", res=1200)
ggplot(May5_6_subset, aes(x = TimeStamp, y = DO_mgL, color = Site)) + 
  scale_color_manual(values = palette_v) +
  theme_pubr() + 
  geom_line() +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ylab("Dissolved Oxygen (mg/L)") + 
  xlab("Time") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme(axis.title=element_text(size=14,face="bold")) +
  theme(axis.text=element_text(size=10)) + 
  theme(legend.text=element_text(size=10,face="bold")) +
  theme(legend.title = element_blank()) +
  #scale_x_datetime(date_breaks = "1 day", date_labels = "%b %d") + 
  geom_hline(yintercept=8, linetype='dotted')
dev.off()
```
