---
title: "EnglishSole_PCA"
author: "Juliana Cornett"
date: "12/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#install packages (tidyverse for data manipulation, and ggpubr, factoextra, and viridis for data visualization/plots)

```{r}
#install.packages("ggpubr")
#install.packages("tidyverse")
#install.packages("factoextra")
#install.packages("viridis")
#install.packages("FactoMineR")
#install.pacakges("dplyr")
#install.packages("patchwork")
```

#load libraries

```{r}
library(ggpubr)
library(tidyverse)
library(factoextra)
library(viridis)
library(FactoMineR)
library(dplyr)
library(patchwork)
```

#read in csv file containing all variables associated with P. vetulus fish from the ventilation trials (ID, Species, DO, weight, length, Hct, Vent, Lac, SOD, and HIF)

```{r}
sole_vars <- read.csv("Englishsole_AllVars.csv")
```

#remove all rows with NA since PCA doesn't work with missing values

```{r}
sole_na <- na.omit(sole_vars)
```

#run PCA on all quantitative variables(columns 7-11), excluding ID, species, DO, length, weight, and ambient ventilation rate

```{r}
sole.pca <- prcomp(sole_na[,c(7:11)], center = TRUE,scale. = TRUE)
summary(sole.pca)
```

#plot relative contribution of variables to principle components (Figure 14 in thesis)

```{r}
fviz_pca_var(sole.pca, col.var = "contrib", gradient.cols = c("#482677FF", "#20A387FF", "#FDE725FF"), 
             repel = TRUE, legend.title = "Contribution") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
  labs(title = NULL, x = "Principal Component 1 (41.3%)", y = "Principal Component 2 (21.6%)") +
  theme(axis.title=element_text(size=14,face="bold")) 
```

#color individuals by DO (Figure 15 in thesis)

```{r}
#jpeg("Sole_indPCA.jpg", width = 7, height = 5, units="in", res=1200)
fviz_pca_ind(sole.pca, label="none", habillage=sole_na$DO, addEllipses=TRUE, 
             ellipse.level=0.95, legend.title = "DO (mg/L)") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  labs(title = NULL, x = "Principal Component 1 (41.3%)", y = "Principal Component 2 (21.6%)") +
  theme(axis.title=element_text(size=14,face="bold")) +
  scale_colour_viridis_d(option = "viridis") +
  scale_fill_viridis_d(option = "viridis")
#dev.off() 
```

#plot with ellipses & variables arrows (figure used in defense)

```{r}
fviz_pca_biplot(sole.pca, label="var", repel = TRUE, alpha.ind=0, mean.point=FALSE, col.var = "black", habillage=sole_na$DO,
               addEllipses=TRUE, ellipse.level=0.95, legend.title = "DO (mg/L)") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
   labs(title = NULL, x = "Principal Component 1 (41.3%)", y = "Principal Component 2 (21.6%)") +
  theme(axis.title=element_text(size=14,face="bold")) +
  scale_colour_viridis_d(option = "viridis") +
  scale_fill_viridis_d(option = "viridis")
```

#PCA with scores

```{r}
princomp(sole_na[,c(7:11)], cor = FALSE, scores = TRUE)
```

#add DO as supplementary qualitative variable ("quali.sup") - DO is column 3

```{r}
res.pca <- PCA(sole_na[,c(7:11)], quali.sup = 3, graph = FALSE)
```

```{r}
print(res.pca)
```

```{r}
ind <- get_pca_ind(res.pca)
```

```{r}
ind$coord[,1]
```

#create new dataframe of individual PC scores with DO levels and coordinates for PC1 and PC2

```{r}
ind_PC <- select(sole_na, c("DO"))
ind_PC$coordPC1 <- ind$coord[,1]
ind_PC$coordPC2 <- ind$coord[,2]
ind_PC
```

#PC1

#create dataframes with mean and variance

```{r}
sole_mPC1 <- group_by(ind_PC, DO) %>% summarize(meanPC1 = mean(coordPC1))
sole_vPC1 <- group_by(ind_PC, DO) %>% summarize(varPC1 = var(coordPC1))
```

#PC2

#create dataframes with mean and variance

```{r}
sole_mPC2 <- group_by(ind_PC, DO) %>% summarize(meanPC2 = mean(coordPC2))
sole_vPC2 <- group_by(ind_PC, DO) %>% summarize(varPC2 = var(coordPC2))
```

#linear models

#PC1

```{r}
sole_mPC1_lm <- lm(ind_PC$coordPC1 ~ ind_PC$DO)
print(summary(sole_mPC1_lm))
```

#variance in PC1

```{r}
sole_vPC1_lm <- lm(sole_vPC1$varPC1 ~ sole_vPC1$DO)
print(summary(sole_vPC1_lm))
```

#PC2

```{r}
sole_mPC2_lm <- lm(ind_PC$coordPC2 ~ ind_PC$DO)
print(summary(sole_mPC2_lm))
```

#variance in PC2

```{r}
sole_vPC2_lm <- lm(sole_vPC2$varPC2 ~ sole_vPC2$DO)
print(summary(sole_vPC2_lm))
```

#combine plots with patchwork (Figure 16 in thesis)

```{r}
#Regression of PC scores on PC1 vs DO

sole_pc1_p1 <- ggplot(ind_PC, aes(DO, coordPC1)) +
  geom_point(fill = "lightskyblue1", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("Individual PC Score (PC1)") + 
  geom_text(aes(x=7, y=3, label= "R^2 == 0.17~~~p == 0.01"), size=5, parse = TRUE)

#Regression in PC1 PC Score Variance vs DO

sole_pc1_p2 <- ggplot(sole_vPC1, aes(DO, varPC1)) + 
  geom_point(fill = "lightskyblue1", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("Variance in Individual PC Score (PC1)") + 
  geom_text(aes(x=7, y=2.5, label= "R^2 == 0.92~~~p == 0.002"), size=5, parse = TRUE)

#Regression of Individual PC scores on PC2 vs DO

sole_pc2_p3 <- ggplot(ind_PC, aes(DO, coordPC2)) +
  geom_point(fill = "lightskyblue1", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("Individual PC Score (PC2)") + 
  geom_text(aes(x=7, y=3, label= "R^2 < 0.001~~~p == 0.86"), size=5, parse = TRUE)

#Regression in PC2 PC Score Variance vs DO

sole_pc2_p4 <- ggplot(sole_vPC2, aes(DO, varPC2)) + 
  geom_point(fill = "lightskyblue1", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("Variance in Individual PC Score (PC2)") + 
  geom_text(aes(x=7, y=2.75, label= "R^2 == 0.15~~~p == 0.45"), size=5, parse = TRUE)

(sole_pc1_p1 | sole_pc1_p2)/
  (sole_pc2_p3 | sole_pc2_p4)
```

