---
title: "SpeckledSanddab_PCA"
author: "Juliana Cornett"
date: "12/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#install packages (tidyverse for data manipulation, and ggpubr, factoextra, and viridis for data visualization/plots)

```{r}
#install.packages("ggpubr")
#install.packages("tidyverse")
#install.packages("factoextra")
#install.packages("viridis")
#install.packages("FactoMineR")
#install.pacakges("dplyr")
#install.packages("patchwork")
#install.packages("outliers")
```

#load libraries

```{r}
library(ggpubr)
library(tidyverse)
library(factoextra)
library(viridis)
library(FactoMineR)
library(dplyr)
library(patchwork)
library(outliers)
```

#read in csv file containing all variables associated with C. stigmaeus fish from the ventilation trials (ID, Species, DO, weight, length, Hct, Vent, Lac, SOD, and HIF)

```{r}
dab_vars <- read.csv("Speckledsanddab_AllVars.csv")
```

#remove all rows with NA since PCA doesn't work with missing values

```{r}
dab_na <- na.omit(dab_vars)
```

#run PCA on all quantitative variables (columns 7-11), excluding ID, species, DO, length, weight, and ambient ventilation rate

```{r}
dab.pca <- prcomp(dab_na[,c(7:11)], center = TRUE,scale. = TRUE)
summary(dab.pca)
```

#plot relative contribution of variables to principle components (Figure 14 in thesis)

```{r}
fviz_pca_var(dab.pca, col.var = "contrib", gradient.cols = c("#482677FF", "#20A387FF", "#FDE725FF"), repel = TRUE, legend.title = "Contribution") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
  labs(title = NULL, x = "Principal Component 1 (25.6%)", y = "Principal Component 2 (24.5%)") +
  theme(axis.title=element_text(size=14,face="bold")) 
```

#color individuals by DO (Figure 15 in thesis)

```{r}
fviz_pca_ind(dab.pca, label="none", habillage=dab_na$DO,
     addEllipses=TRUE, ellipse.level=0.95, legend.title = "DO (mg/L)") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
   labs(title = NULL, x = "Principal Component 1 (25.6%)", y = "Principal Component 2 (24.5%)") +
  theme(axis.title=element_text(size=14,face="bold")) +
  scale_colour_viridis_d(option = "viridis") +
  scale_fill_viridis_d(option = "viridis") 
```

#plot with ellipses & variables arrows (figure used in defense)

```{r}
fviz_pca_biplot(dab.pca, label="var", repel=TRUE, alpha.ind=0, mean.point=FALSE, col.var = "black", habillage=dab_na$DO,
               addEllipses=TRUE, ellipse.level=0.95, legend.title = "DO (mg/L)") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
   labs(title = NULL, x = "Principal Component 1 (25.6%)", y = "Principal Component 2 (24.5%)") +
  theme(axis.title=element_text(size=14,face="bold")) +
  scale_colour_viridis_d(option = "viridis") +
  scale_fill_viridis_d(option = "viridis")
```

#PCA with scores

```{r}
princomp(dab_na[,c(7:11)], cor = FALSE, scores = TRUE)
```

#add DO as supplementary qualitative variable ("quali.sup") - DO is column 3

```{r}
res.pca <- PCA(dab_na[,c(7:11)], quali.sup = 3, graph = FALSE)
```

```{r}
print(res.pca)
```

```{r}
ind <- get_pca_ind(res.pca)
```

```{r}
ind$coord[,1]
```

#create new dataframe of individual PC scores with DO levels and coordinates for PC1 and PC2

```{r}
ind_PC <- select(dab_na, c("DO"))
ind_PC$coordPC1 <- ind$coord[,1]
ind_PC$coordPC2 <- ind$coord[,2]
ind_PC
```
#use Grubbs test to determine if high value for PC2 at 1.5 mg/L O2 is outlier

```{r}
grubbs_test_pc2 <- grubbs.test(ind_PC$coordPC2)
grubbs_test_pc2
```

#create dataframe with PC2 scores only without outlier

```{r}
#select only DO and PC2 scores
ind_PC2 <- select(ind_PC, c("DO", "coordPC2"))
#remove row with outlier (42nd row)
indPC2_noOutlier <- ind_PC2[-c(42), ]
indPC2_noOutlier
```

#PC1

#create dataframes with mean and variance

```{r}
dab_mPC1 <- group_by(ind_PC, DO) %>% summarize(meanPC1 = mean(coordPC1))
dab_vPC1 <- group_by(ind_PC, DO) %>% summarize(varPC1 = var(coordPC1))
```

#PC2

#create dataframes with mean and variance

```{r}
dab_mPC2 <- group_by(indPC2_noOutlier, DO) %>% summarize(meanPC2 = mean(coordPC2))
dab_vPC2 <- group_by(indPC2_noOutlier, DO) %>% summarize(varPC2 = var(coordPC2))
```

#linear models

#PC1

```{r}
dab_mPC1_lm <- lm(ind_PC$coordPC1 ~ ind_PC$DO)
print(summary(dab_mPC1_lm))
```

#variance in PC1

```{r}
dab_vPC1_lm <- lm(dab_vPC1$varPC1 ~ dab_vPC1$DO)
print(summary(dab_vPC1_lm))
```

#PC2

```{r}
dab_mPC2_lm <- lm(indPC2_noOutlier$coordPC2 ~ indPC2_noOutlier$DO)
print(summary(dab_mPC2_lm))
```

#variance in PC2

```{r}
dab_vPC2_lm <- lm(dab_vPC2$varPC2 ~ dab_vPC2$DO)
print(summary(dab_vPC2_lm))
```

#try nonlinear regression for PC2 scores & variance for speckled sanddabs

#polynomial regression for PC2 scores

```{r}
lm(coordPC2 ~ poly(DO, 2, raw = TRUE), data = indPC2_noOutlier) %>%
  summary()
```

#plot nonlinear for PC2 scores

```{r}
ggplot(indPC2_noOutlier, aes(DO, coordPC2) ) +
  geom_point(fill = "lightgoldenrod2", shape = 21, size = 5) +
  theme_pubr() +
  stat_smooth(method='lm', color="black", se = FALSE, formula = y ~ poly(x, 2)) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("Individual PC Score (PC2)") +
  geom_text(aes(x=2.5, y=2, label= "R^2 == 0.32~~~p < 0.001 "), size=5, parse = TRUE)
```


#polynomial regression for PC2 score variance

```{r}
lm(varPC2 ~ poly(DO, 2, raw = TRUE), data = dab_vPC2) %>%
  summary()
```

#plot nonlinear for PC2 score variance

```{r}
ggplot(dab_vPC2, aes(DO, varPC2) ) +
  geom_point(fill = "lightgoldenrod2", shape = 21, size = 5) +
  theme_pubr() +
  stat_smooth(method='lm', color="black", se = FALSE, formula = y ~ poly(x, 2)) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("Variance in Individual PC Score (PC2)") +
  geom_text(aes(x=2.5, y=0.55, label = "R^2 == 0.54~~~p == 0.31"), size = 5, parse = TRUE)
```

#combine plots with patchwork (Figure 17 in thesis)

```{r}
#Regression of PC scores on PC1 vs DO

dab_pc1_p1 <- ggplot(ind_PC, aes(DO, coordPC1)) +
  geom_point(fill = "lightgoldenrod2", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("Individual PC Score (PC1)") + 
  geom_text(aes(x=7, y=2.75, label= "R^2 < 0.001~~~p == 0.98"), size=5, parse = TRUE)

#Regression in PC1 PC Score Variance vs DO

dab_pc1_p2 <- ggplot(dab_vPC1, aes(DO, varPC1)) + 
  geom_point(fill = "lightgoldenrod2", shape = 21, size = 5) +
  theme_pubr() +
  geom_smooth(method = "lm", color ="black", se = FALSE) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("Variance in Individual PC Score (PC1)") +
  geom_text(aes(x=7, y=2.5, label= "R^2 == 0.25~~~p == 0.31"), size=5, parse = TRUE)

#Nonlinear Regression of Individual PC scores on PC2 vs DO

dab_pc2_p3 <- ggplot(indPC2_noOutlier, aes(DO, coordPC2) ) +
  geom_point(fill = "lightgoldenrod2", shape = 21, size = 5) +
  theme_pubr() +
  stat_smooth(method='lm', color="black", se = FALSE, formula = y ~ poly(x, 2)) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("Individual PC Score (PC2)") +
  geom_text(aes(x=7, y=2, label= "R^2 == 0.32~~~p < 0.001 "), size=5, parse = TRUE)

#Regression in PC2 PC Score Variance vs DO

dab_pc2_p4 <- ggplot(dab_vPC2, aes(DO, varPC2) ) +
  geom_point(fill = "lightgoldenrod2", shape = 21, size = 5) +
  theme_pubr() +
  stat_smooth(method='lm', color="black", se = FALSE, formula = y ~ poly(x, 2)) +
  theme(axis.text = element_text(size=14), axis.title = element_text(size=18, face="bold")) +
  xlab("DO (mg/L)") +
  ylab("Variance in Individual PC Score (PC2)") +
  geom_text(aes(x=7, y=0.575, label = "R^2 == 0.54~~~p == 0.31"), size = 5, parse = TRUE)

(dab_pc1_p1 | dab_pc1_p2)/
  (dab_pc2_p3 | dab_pc2_p4)
```
